---
title: "Birthday"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Birthday}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(MATH4753FALLKistler)
```

## Section 1: LaTex and proving:
\[
\textbf{What are we finding?}\quad
\text{The chance that in a group of } n \text{ people, at least two share a birthday.}
\]

\[
\text{Assume birthdays are independent and spread evenly across 365 days.}
\]

\[
\text{It’s easier to find the opposite: everyone has a different birthday.}
\]
\[
\Pr(\ge 2 \text{ share}) \;=\; 1 - \Pr(\text{all birthdays are different}).
\]

\[
\text{Build it step by step: the first person can be on any day (prob }1\text{).}
\]
\[
\text{The second must avoid that day } \Big(\tfrac{364}{365}\Big),\;
\text{the third must avoid two days } \Big(\tfrac{363}{365}\Big),\; \ldots
\]
\[
\Rightarrow\quad
\Pr(\text{all different})
= \frac{365}{365}\cdot\frac{364}{365}\cdot\frac{363}{365}\cdots
\frac{365-n+1}{365}.
\]

\[
\text{Same thing, just written with factorials:}\quad
\Pr(\text{all different}) \;=\; \frac{365!}{(365-n)!\,365^n}.
\]

\[
\boxed{\;\Pr(\ge 2 \text{ share}) \;=\; 1 - \dfrac{365!}{(365-n)!\,365^n}\;}
\]

\[
\text{\small (Why the fraction? There are } 365^n \text{ total birthday patterns, and }
365\cdot 364\cdots(365-n+1)=\frac{365!}{(365-n)!} \text{ of them have no repeats.)}
\]





## section 2: Taking this into R now

We calculate $P(\ge 1\ \text{shared birthday})$ for a group of size $x$ with `birthday()`.

### The R code in question:
````{r}
#' Title
#'
#' @param x an integer that is the size of the group to sample
#'
#' @returns a number which is the probability of at least one shared birthday
#' @export
#' @examples
#' birthday(10)
#' birthday(40)
birthday <- function(x)
{
  # log P(all birthdays distinct) = log(365! / ((365-x)! * 365^x))
  L <- lchoose(365, x) + lfactorial(x) - x * log(365)
  1 - exp(L)  # complement => P(at least one shared birthday)
}

### Examples
````{r}
birthday(20:25)
````
